NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PS lipids

### 2018-02-01 Manuscript/manuscriptPS.tex
Add Ca density plots and some updates

The headgroup order parameter changes of POPC and POPS from POPC:POPS (5:1) bilayers
as a function of Ca$2+$ concentration from different
simulations and experiments \cite{roux90} are shown in Fig. \ref{changesWITHCaClPG}.
As shown previously, the changes of PC headgroup order parameters can be
used to measure the amount of bound ions to the bilayer according to the
electrometer concept \cite{akutsu81,seelig87,catte16}.
This would suggest that the Ca2+ binding is overestimated in
MacRog simulation and underestimated in CHARMM36 simulation.
It should be however noted, that MacRog simulation with low
concetrations gives a good agreemet with experiments.
The result from CHARMM simulation is somewhat surprising because
previous study suggested that Ca2+ ions overbind to pure PC lipid
bilayers \cite{catte16}.
\todo{The role of counterions and equilibration should be somehow checked
and discussed.}
\begin{figure*}[]
  \centering
  \includegraphics[width=18cm]{../Figs/CHANGESwithCaClPS.eps}
  \caption{\label{changesWITHCaClPG}
    POPC (left) and POPS (right) headgroup order parameters in lipid bilayer with POPC:POPS (5:1) mixture
    as a function CaCl$_2$ concentration from experiments \cite{roux90} and different simulations.
  }
\end{figure*}

>>>>

The headgroup order parameter changes of POPC and POPS from POPC:POPS (5:1) bilayers
as a function of Ca$2+$ concentration from different
simulations and experiments \cite{roux90} are shown in Fig. \ref{changesWITHCaClPG}.
As shown previously, the changes of PC headgroup order parameters can be
used to measure the amount of bound ions to the bilayer according to the
electrometer concept \cite{akutsu81,seelig87,catte16}.
This would suggest that the Ca2+ binding is overestimated in
MacRog simulation and underestimated in CHARMM36 simulation.
It should be however noted, that MacRog simulation with low
concentrations gives a good agreemet with experiments.
The result from CHARMM simulation is somewhat surprising because
previous study suggested that Ca2+ ions overbind to pure PC lipid
bilayers \cite{catte16}.
\todo{The role of counterions and equilibration should be somehow checked
  and discussed.}
The density profiles of Ca2+ from the simulations are shown in Fig. \ref{CAdensPCPSmixture}.
Overall, it seems that calcium clearly overbinds in MacRog simulations
as expected based on previous results \cite{catte16}.
However, the calcium binding in CHARMM36 simulations seems to
surprisingly low. Thi
\begin{figure*}[]
  \centering
  \includegraphics[width=18cm]{../Figs/CHANGESwithCaClPS.eps}
  \caption{\label{changesWITHCaClPG}
    POPC (left) and POPS (right) headgroup order parameters in lipid bilayer with POPC:POPS (5:1) mixture
    as a function CaCl$_2$ concentration from experiments \cite{roux90} and different simulations.
  }
\end{figure*}
\begin{figure}[]
  \centering
  \includegraphics[width=9cm]{../Figs/CAdensPCPSmixture.eps}
  \caption{\label{CAdensPCPSmixture}
    Ca2+ density profiles from simulations.
  }
  \todo{Ca2+ binding in CHARMM is lower than to neutral membrane. The reason should be found out (probably equilibration).}
  \todo{These are now mass densities, numbers would be probably better.}
  \todo{Counterions should be also included.}
  \todo{Not all the data from MacRog is included.}
\end{figure}
